---
output:
  pdf_document: 
    toc: true
    number_sections: true
    keep_tex:  true
  html_document: 
      collapsed: false
      smooth_scroll: false
      number_sections: true
      theme: united
      highlight: tango
  output:
    md_document:
      variant: markdown_github
author: "Lea Lai"
date: "February 23, 2019"
title: "Get start with R"
---


# A start: Get use to R 
(Partially credit to Nicole Kelbick, PhD. and introduction to R https://cran.r-project.org/doc/manuals/r-release/R-intro.pdf)

Don't afriad to use R. It can be very simple. You can start by open R and type in only one line, and it will work. Try the following:

(The ones with ** are frequently used)

## Common used operation or funtions

### ** ":"
The operator ':' generates a sequence of integers. 
```{r}
1:10
10:1
```

### "<-" or "="
You can assign values to variables using '<-' OR '='.
```{r}
x <- 5
x
x=2
x
```

### "+" "-" "*" "/" "%%"
These are basic arithmetic operations 
```{r}
x+5
x*5
x/5
x%%5 #give the remainder of x
```

### ** "seq(from,to,spacing)" 
The 'seq' function generates a sequence of numbers with a specified spacing.\
seq( from,to,spacing)
```{r}
xn <- seq(1,10,.1)
xn
seq(1,10,length.out = 20) #use length.out to specify how many you need within the range

```

### "rev"
The 'rev' function reverses values of argument.

```{r}
yn <- rev(xn)
yn
```


### ** "c(elem1,elem2)"
The operator 'c'combines different elements into a vector
```{r}
c(1,2)
c("1",2) #the same as c("1","2"), they are all stored as strings. 
```

### ** rep(arg1,n)
'rep(arg1, n)' repeats the first argument (arg1) n times
```{r}
rep(2,7)
y <- c(1, 3, 5.5, rep(2,7))
y

rep(c(1,3),3) # repeat 1,3 for 3 times

rep(seq(1,3),2:4) # repeat 1,2,3 correspondingly for 2,3,4 times
```

### Type casting: as.numeric and etc.
Change string to number; or change number to string 
```{r}
as.numeric("1") #when you add  " " , the content in the double quotation marks become strings
as.character(1)

```

### ";"
The operator ";" can be used as a seperation for each command when writing on the same line
```{r}
print(x);print(y)
```

### "mean";"var";"sd","median"
```{r}
x=1:10;x
mean(x) #Get average
median(x) #Get median
var(x) #Get variation
sd(x) #Get Standard deviation
```

### "paste(elm1,elm2,sep)"
the paste(element1,element2,sep="") function combines elements into strings
```{r}
paste("Day",1:10,sep="")
paste("Day",1:10,sep="_")
paste(c("X","Y"), 1:10, sep="")
```



## Logical expressions 

### or "|" "||"


When comparing single value, you may use  "|" or "||"  
```{r}
x=2;y=3
y > 0 | x >= 3
y > 0 || x >= 3
```
When comparing a vector, you use "|" to gain results of comparison by array 
```{r}
a=1:3;b=2:4
a>b | a==b
```
"||" gives only a single logic value when comparing a vetor  
```{r}
a>b || a==b
```

### and "&" "&&"
```{r}
a>b & a==b
y > 0 && x >= 3
```


## About functions
### Find help "?"
Find help/example/instruction for function, add a "?" before the function name:
```{r, eval=F, echo=T}
?print
```

```{r, attach a png, echo=FALSE,out.width='50%'}
knitr::include_graphics(paste(getwd(),"/pics/print_exp.png",sep=""),dpi=500)
```


Use single quotes to get help on operators
```{r, eval=F, echo=T}
?`:`
```


These helping information in the picture above will show up on the sidebar of R studio



### "args": View arguments of function 
To view the list of possible arguments a function can have use 'args'
```{r}
args(png) #png is a function to export graph as png in your computer

```
### View the whole function
To see the whole function: type in the function name without "()" followed 
```{r}
var
```


### Get/set Working dictionary: "getwd()" "setwd()"
Get current working dictionary:
```{r}
getwd()
```

Set current working dictionary:
```{r,eval=F, echo=T}
setwd("C:/Users/naszh/Desktop")
```

As we started, (e.x.: a=c(1,2,...)) is a way to combine elelments and create vectors. There also are other ways to create vectors:

## vector or array

Create vector or array
```{r}
x=vector()
x[3]=10
x

y=array()
y[4]=1
y
```
Difference between array and vector is that array can have more dimensions than vector:

```{r}
array(dim=c(1,2))
```
dim stands for dimension at here.

### Assign names for a vector or array
Vectors can have names for each element, and array can have column names and row names
```{r}
x=1:10 #x become a vector
names(x)=paste("X",1:10,sep="")
x

y=array(2:3,dim=c(1,2))
colnames(y)=c("col1","col2") #define column names
rownames(y)="row1" #define row names
y
```
### Call elements in vector or array
call by names (use x and y value from above)
```{r}
x["X1"] #don't forget to add " " on the name inside []
y["row1","col1"]
```
Call by index number
```{r}
#vector:
x=1:10
x
x[1]
x[1:2]

#array:
y=array(1:20,dim=c(4,5))
y
y[2,3] #row2, column 3
y[2, ] #present row 2


```


### Select only certain things in the array
```{r}
x=1:10
x[x>5]
```
"x>5" is a logical statement and give an array of logical vlaues:
```{r}
x>5
```
This logical value array can be put into  

##Dataframe

create dataframe:
```{r}
d = data.frame("Feature1" = c(1,10,rep(12,2)))
d
```

Call a column:
```{r}
d$Feature1
```

Add a column:
```{r}
d = cbind(d, "Feature2" = c(10,29,9,2)) #cbind stand for column bind;
    #similarly, rbind is for row bind. 
d
```




##Ploting

### "plot"
You can plot data using the function 'plot'
```{r}
x = 1:10;y=2:11
plot(x,y)

```

### Export plot: "pdf", "png"
Export plot  as a pdf (or other formats).
```{r}
pdf(file="homework1_plot2.pdf", height=3, width=3)
plot(y,x)
plot(x,x)
dev.off()
  #Run these functions together
  #the first commend "pdf" startsthe graphics device to pdf, 
  #and the following graphics would be produced in the pdf  
  #When finish plotting, use dev,off to turns off the connection to the graphical device.
  #The file will show up in whatever your current working directory is.
```


### Change font size: "plot(... , lwd= )"
Use larger font for axis labels
```{r,out.width="50%"}
plot(x, y, type='l', lwd=8)
plot(x, y, type='l', lwd=3)
```

### Change font color: "plot(... , col= )"
```{r,out.width="50%"}
plot(x, y, type='l', lwd=8, col='blue')
```

### Change type of plot: "plot(..., type= )" 
```{r,out.width="50%"}
plot(x, y, type='l') # "l" for lines
plot(x, y, type='p') # "p" for points
plot(x, y, type='b') # "b" for both
```
More types usage see "?plot"

### Change title/ lab names
```{r,out.width="50%"}
plot(x, y, type='l',xlab="Time",ylab="Grade",main="Time-Grade",sub="Plot 1")
```

